---
title: "Exploratory Analysis of the data"
author: "Yash Amonkar"
date: "February 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(dplyr)
```

###Data Input

The variables selected are the FIPS code, the per capita medicine and bottles sales, week end along with indicators for violations in the first and second week. 

```{r}
input_data <- read.csv("cnty_wk_V2_FULLPANEL.csv")
cleaned_data <- input_data %>% 
        select(county_FIPS, week_end, dum_1stwk_viol, dum_2ndwk_viol, defl_medsales_pc, defl_bottlesales_pc)
head(cleaned_data)

```

##Creating the Dataset for violation for first and second weeks

The Na values are converted to zero.

```{r}
cleaned_data$dum_1stwk_viol[is.na(cleaned_data$dum_1stwk_viol)] <- 0  
cleaned_data$dum_2ndwk_viol[is.na(cleaned_data$dum_2ndwk_viol)] <- 0
cleaned_data$dum_2ndwk_viol <- cleaned_data$dum_1stwk_viol + cleaned_data$dum_2ndwk_viol
colSums(is.na(cleaned_data)) #Check for NA Values
head(cleaned_data)
```

## Selecting counties with complete Med Sales Data

We have medicine sales data for select counties. We now select and include the counties for which we have medicine sales data. Note:- The counties which have medicine sales data also have data for sales of bottled water.

```{r, warning=FALSE}
total_med_sales_per_county <- cleaned_data %>% 
                  group_by(county_FIPS) %>%
                  summarise(meds = sum(defl_medsales_pc))

total_med_sales_per_county <- total_med_sales_per_county %>% 
                  filter(meds > 0)
head(total_med_sales_per_county)
```

The file above contains the sum of total medicine sales. Values which correspond to zero, are counties which do not have medicine sales were removed. 

##Merging

```{r}
Merged_dataset <- merge(cleaned_data,total_med_sales_per_county, by = "county_FIPS")
Merged_dataset$meds <- NULL #This is the total med sales for 10 years and is not needed. 
colSums(is.na(Merged_dataset)) # Check for NA values
head(Merged_dataset)
```

##Saving the Dataset

The dataset which is cleaned can be used directly in the Stan-Regression models. We save the output in a text format. 

```{r}
write.table(Merged_dataset, "Dataset for Regression.txt", sep = " ")
```

